\vspace{-1.3cm} \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration

### 2016-03-29 paper/paper.tex
diagramme...

@@

\tikzset{
    block/.style={
           rectangle,
           rounded corners,
           draw=black, very thick,
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
}

@@

\ \\ 
\begin{tikzpicture}[auto]%node distance=0.5cm and 1.8cm]
	\node[block] (prm) {
		\begin{tabular}{c}
		\textbf{PRMSolver}\\
		generate nodes\\
		store graph
		\end{tabular}
	};
	\node [block, right=of prm] (config) {
		\begin{tabular}{c}
		\textbf{Configpace}\\
		indicator fct. \\
		interface
		\end{tabular}
	};
	\node [block, right=of config] (robo) {
		\begin{tabular}{c}
		\textbf{RobotConfigspace}\\
		robot kinematics\\
		collision detection
		\end{tabular}
	};
	\node [block, right=of robo] (gpu) {\textbf{GPU}};

	\path[-stealth,thick, shift left=1.0ex]	(prm) edge (config) ;
	\path[-stealth,thick, shift left=1.0ex]	(config) edge  (prm) ;
	\path[-stealth,thick, shift left=1.0ex]	(config) edge (robo);
	\path[-stealth,thick, shift left=1.0ex]	(robo) edge  (config);
	\path[-,thick]					(robo) edge (gpu);
\end{tikzpicture}
\ \\

@@

\section{Implementation of an Indicator Function for Robot Arms}\label{indicator}.

@@

To calculate the intersection between two polytopes, the Chung-Wang algorithm only needs their vertices 
and the data which vertices are connected by an edge. 
For this the adjacency matrix of the edge graph is stored compactly in the crs format.
\begin{lstlisting}
struct polytope4{
    float4* vertices;	//length n
    int n;
    //!edges in crs format
    int* dsp;	//length n
    int* cnt;	//length n
    int* dest;	//length m
    int m;
};
\end{lstlisting}
It means, that for $i=0,...,n-1$ the vertices $dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.

- whole list of polytopes stored compactly in polytope4data on gpu

- maybe describe interface polytope <-> polytope4data <-> polytope4

@@

@@

PRMSolver 1 <-> Configspace object 1 <-> GPU 1

.

.

.

PRMSolver n <-> Configspace object 2 <-> GPU 1

@@

- definitions as pseudocode

version 1:

	- each process generates own nodes -> exchange them, synchronous gpu kernel

version 2:

@@

version 3:

@@

@@

>>>>

\usepackage{graphicx}

@@

\tikzset{
    block/.style={
           rectangle,
           rounded corners,
           draw=black, very thick,
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
     block2/.style={
           %rectangle,
           %rounded corners,
           fill=black!20,
           %draw=black, very thick,
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
     frameless/.style={
           %rectangle,
           %rounded corners,
           %draw=black, very thick,
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
}

@@

%\ \\
\begin{center} 
\begin{tikzpicture}[auto]%node distance=0.5cm and 1.8cm]
	\node[block] (prm) {
		\begin{tabular}{c}
		\textbf{PRMSolver}\\
		generate nodes\\
		store graph
		\end{tabular}
	};
	\node [block, right=of prm] (config) {
		\begin{tabular}{c}
		\textbf{Configpace}\\
		indicator fct. \\
		interface
		\end{tabular}
	};
	\node [block, right=of config] (robo) {
		\begin{tabular}{c}
		\textbf{RobotConfigspace}\\
		robot kinematics\\
		collision detection
		\end{tabular}
	};
	\node [block, right=of robo] (gpu) {\textbf{GPU}};

	\path[-stealth,thick, shift left=1.0ex]	(prm) edge (config) ;
	\path[-stealth,thick, shift left=1.0ex]	(config) edge  (prm) ;
	\path[-stealth,thick, shift left=1.0ex]	(config) edge (robo);
	\path[-stealth,thick, shift left=1.0ex]	(robo) edge  (config);
	\path[-,thick]					(robo) edge (gpu);
\end{tikzpicture}
\end{center}
%\ \\

@@

\section{Implementation of an Indicator Function for Robot Arms}\label{indicator}

@@

To calculate the intersection between two convex polytopes, the Chung-Wang algorithm only needs their vertices 
and the data which vertices are connected by an edge. 
For this the adjacency matrix of the edge graph is stored compactly in the crs format.
\begin{lstlisting}
struct polytope4{
    float4* vertices;	//length n
    int n;
    //!edges in crs format
    int* dsp;	//length n
    int* cnt;	//length n
    int* dest;	//length m
    int m;
};
\end{lstlisting}
It means, that for $i=0,...,n-1$ the vertices $dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
The polytopes used in this projects are given by the user as convex hull of a list of vertices. However, to determine the edge graph only from the vertices is a nontrivial task, which is done by Matlab here using provided high level functions.

@@

\subsection{Storage of the Graph}

\subsection{Parallelization Models}

@@

@@

- mpi

- every process has its own GPU

\subsubsection*{Version 1}

In the first approach, that was evaluated, every process generates independantly from each other new nodes.
Then the possible neighbors are determined and the connections, which have to be tested.
The indicator function is called and returnes the edges.
After this the graph must be synchronized on every processes, which leads to an alltoall communication.
As in this graph building does not need the edge data, it is sufficient to exchange only the nodes in this phase.

	\begin{center}
\begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
	\node[block, minimum width=3cm] (prm) {
		\begin{tabular}{c}
		\textbf{PRMSolver 1}
		\end{tabular}
	};

	\node[frameless, below=of prm, minimum width=3cm] (space) {$\vdots$};

	\node[block, below=of space, minimum width=3cm] (prm2) {
		\begin{tabular}{c}
		\textbf{PRMSolver n}
		\end{tabular}
	};

	\node [block, right=of prm] (config) {
		\begin{tabular}{c}
		\textbf{Configpace 1}
		\end{tabular}
	};
	\node[frameless, below=of config] (space2) {$\vdots$};
	\node [block, below=of space2] (config2) {
		\begin{tabular}{c}
		\textbf{Configpace n}
		\end{tabular}
	};

	\node [block, right=of config] (gpu) {\textbf{GPU 1}};
	\node [block, right=of config2] (gpu2) {\textbf{GPU n}};

	\path[-latex',thick, shift left=1.0ex]	(prm) edge (config) 
								(config) edge (prm);
	\path[-,thick]					(config) edge (gpu);

	\path[-latex',thick, shift left=1.0ex]	(prm2) edge (config2) 
								(config2) edge (prm2) ;
	\path[-,thick]					(config2) edge (gpu2);

	\path[-latex',thick, shift left=1.0ex]	(prm) edge (space)
								(space) edge (prm)
								(prm2) edge (space)
								(space) edge (prm2) ;

\end{tikzpicture}
\end{center}

\subsubsection*{Version 2}
The second approach

@@

	\begin{center}
\begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
	\node[block, minimum width=3cm] (prm) {
		\begin{tabular}{c}
		\textbf{PRMSolver 1}
		\end{tabular}
	};

	\node[frameless, below=of prm, minimum width=3cm] (space) {$\vdots$};

	\node[block, below=of space, minimum width=3cm] (prm2) {
		\begin{tabular}{c}
		\textbf{PRMSolver n}
		\end{tabular}
	};

	\node [block, right=of prm] (config) {
		\begin{tabular}{c}
		\textbf{Configpace 1}
		\end{tabular}
	};
	\node[frameless, below=of config] (space2) {$\vdots$};
	\node [block, below=of space2] (config2) {
		\begin{tabular}{c}
		\textbf{Configpace n}
		\end{tabular}
	};

	\node [block, right=of config] (gpu) {\textbf{GPU 1}};
	\node [block, right=of config2] (gpu2) {\textbf{GPU n}};

	\path[-latex',thick, shift left=1.5ex]	(prm) edge (config) ;
	\path[-,thick]					(config) edge (gpu);

	\path[-latex',thick, shift right=1.5ex]	(prm2) edge (config2) ;
	\path[-,thick]					(config2) edge (gpu2);

	\path[-latex',thick]				(config.west) edge  (prm.east) 
								(config.west) edge (space.east)
								(config.west) edge (prm2.east)
								(config2.west) edge  (prm.east) 
								(config2.west) edge (space.east)
								(config2.west) edge (prm2.east);

\end{tikzpicture}
\end{center}

\subsubsection*{Version 3}

@@

Stateflow for each PRMSolver instance:

\begin{center}
\begin{tikzpicture}
	\tikzset{mywidth/.style={minimum width=2.5cm},};
	\node[block, mywidth] (l1) {\textbf{Worker 1}};
	\node[frameless, mywidth, below=of l1, minimum height=1.5cm] (l2) {$\vdots$};
	\node[block, mywidth, below=of l2] (l3) 	{\begin{tabular}{c}
										generate\\ nodes, \\
										start kernel
									\end{tabular}};
	\node[frameless, mywidth, below=of l3, minimum height=2cm] (l6) {};
	\node[frameless, mywidth, below=of l6] (l7) {};
	\node[block, mywidth, below=of l7] (l11)	{\begin{tabular}{c}
										recieve\\ result, \\
										store nodes \\ 
										and edges
									\end{tabular}};
	\node[frameless, mywidth, below=of l11] (l12) {$\vdots$};

	\node[block, mywidth, right=of l1] (r1) {\textbf{Worker 2}};
	\node[frameless, mywidth, right=of l2, minimum height=1.5cm] (r2) {$\vdots$};
	\node[frameless, mywidth, right=of l3] (r3) {};
	\node[block, mywidth, right=of l6] (r6)	{\begin{tabular}{c}
										recieve\\ result, \\
										store nodes \\
										and edges
									\end{tabular}};
	\node[block, mywidth, right=of l7] (r7) 	{\begin{tabular}{c}
										generate\\ nodes, \\
										start kernel
									\end{tabular}};
	\node[frameless, mywidth, right=of l12] (r12) {$\vdots$};

	\node[frameless, left=of l1] (g1) {};
	\node[block, below=of g1, minimum height=14cm] (graph) {\rotatebox{90}{\textbf{Graph}}};

	%gpu column
	\tikzset{gpuwidth/.style={minimum width=0.8cm},};

	\node[block, right=2cm of r1, minimum width=2.5cm] {\textbf{GPU}};
	\node[frameless, right=2cm of r2, gpuwidth] (topa) {$\vdots$};
	\node[frameless, right=of topa, gpuwidth] (topb) {$\vdots$};
	\node[block2, below=0.7cm of topa, gpuwidth, minimum height=4.3cm] (stream1a) {\rotatebox{90}{kernel}};
	\node[block2, below=2.3cm of stream1a, gpuwidth, minimum height=4.2cm] (stream1b) {\rotatebox{90}{kernel}};
	\node[frameless, below=0.3cm of stream1b, gpuwidth] (bota) {$\vdots$};

	\node[block2, below=2.2cm of topb, gpuwidth, minimum height=7.6cm] (stream2) {\rotatebox{90}{kernel}};
	\node[frameless, right=of bota, gpuwidth] (botb) {$\vdots$};

	%EDGES

	\path[-latex',thick]	(graph.east |- l3.west) edge (l3.west)
					(l11.west) edge (graph.east |- l11.west);
	\path[-latex',thick]	(graph.east |- r7.west) edge (r7.west)
					(r6.west) edge (graph.east |- r6.west);
	\path[dotted,thick]	(l2) edge (l3)
					(l3) edge (l11)
					(l11) edge (l12)
					(r2) edge (r6)
					(r6) edge (r7)
					(r7) edge (r12);

	\path[-latex',thick]	(r6.east -| stream1b.west) edge (r6.east);
	\path[-latex',thick]	(r7.east) edge (r7.east -| stream1b.west);

	\path[-,thick]		(l3.east) edge (l3.east -| stream1a.west);
	\path[-latex',thick]	(stream1a.east |- l3.east) edge (l3.east -| stream2.west);
	\path[latex'-,thick]		(l11.east) edge (l11.east -| stream1b.west);
	\path[-,thick]		(stream1b.east |- l11.east) edge (l11.east -| stream2.west);

\end{tikzpicture}
\end{center}

@@

- difficult to test because
	- not deterministic
	- dependant on density of nodes: how many neighbours have to be tested per connection
